---
title: "exercise_3"
author: "Gioele Pinana"
date: '2022-05-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task and Inputs

In this task we are using the method proposed in Laube and Purves (2011) about 
segmenting trajectories. The task will be to understand this implementation and 
apply it on Caro, with a different sampling interval.

## libraries 

```{r message=FALSE, warning=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(SimilarityMeasures)
library(tidyr)
```

## Task 1: Segmentation   

Importing dataset 
```{r}
caro <- read_csv("caro60.csv")
```

The sampling interval for this dataset is 1 minute. We use a temporal window 
of 6 minutes. We need to calculate the following Euclidean distances 
(pos representing single location):

1. pos[n-3] to pos[n]
2. pos[n-2] to pos[n]
3. pos[n-1] to pos[n]
4. pos[n] to pos[n+1]
5. pos[n] to pos[n+2]
6. pos[n] to pos[n+3]

```{r}
caro2 <- caro %>%
  mutate(
    nMinus3 = sqrt((lag(E,3)-E)^2+(lag(N,3)-N)^2),   # distance to pos -3 minutes
    nMinus2 = sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),   # distance to pos -2 minutes
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),   # distance to pos -1 minutes
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2), # distance to pos +1 mintues
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2), # distance to pos +2 minutes
    nPlus3  = sqrt((E-lead(E,3))^2+(N-lead(N,3))^2)  # distance to pos +3 minutes
  )
```

## Task 2: Specify and apply threshold d

```{r}
caro3 <- caro2 %>%
  rowwise() %>%
  mutate(
    stepMean = mean(c(nMinus3, nMinus2, nMinus1,nPlus1,nPlus2, nPlus3))
  ) %>%
  ungroup()
```

In the last code line we calculated the mean distance of n Minus 3, nMinus2, 
nMinus1, nPlus1, nPlus2 and nPlus3 for each row. Now we can determine if an 
animal is moving or not by specifying a threshold distance on stepMean. 
For this task we use the mean value as a threshold: Positions with distances 
below this value are considered static

```{r}
caro4 <- caro3 %>% 
  ungroup() %>%
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE))

caro_filter <- caro4 %>%
  filter(!static)
```

## Task 3: Visualize segmented trajectories

Now visualize the segmented trajectory spatially
```{r}

ggplot(caro4, aes(E, N, colour = static)) +
  geom_path() +
  geom_point() +
  coord_equal()

ggplot(caro_filter, aes(E, N)) +
  geom_path() +
  geom_point() +
  coord_equal()
```

## Task 4: Segment-based analysis

```{r}
rle_id <- function(vec){
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times=x))
}

caro5 <- caro4 %>%
  mutate(segment_id = rle_id(static))%>%
  filter(!static) %>%
  group_by(segment_id) %>%
  mutate(duration = n()) %>%
  filter(duration>6)
```


```{r figures-side, fig.show="hold", out.width="50%"}
par(mfrow=c(1,2))

ggplot(caro5, aes(E, N, colour = segment_id)) +
  geom_path() +
  geom_point() +
  coord_equal() +
  labs(title = "All segments")

ggplot(caro5, aes(E, N, colour = segment_id)) +
  geom_path() +
  geom_point() +
  coord_equal() +
  labs(title = "Segments longer than 5 minutes")
```

## Task 5: Similarity measures

Importing dataset 
```{r}
pedestrians <- read_csv("pedestrian.csv")
```

Visual comparision of the 6 trajectories
```{r}
ggplot(pedestrians, aes(N,E, col = as.factor(TrajID)))+
  geom_point() +
  geom_path() +
  facet_wrap(~TrajID)
```

## Task 6: Calculate similarity

First I create the matrices 
```{r}
pedestrians1 <- pedestrians %>%
  filter(TrajID == 1) %>%
  select(E,N)%>%
  as.matrix() 

pedestrians2 <- pedestrians %>%
  filter(TrajID == 2) %>%
  select(E,N) %>%
  as.matrix() 

pedestrians3 <- pedestrians %>%
  filter(TrajID == 3) %>%
  select(E,N) %>%
  as.matrix() 

pedestrians4 <- pedestrians %>%
  filter(TrajID == 4) %>%
  select(E,N) %>%
  as.matrix() 

pedestrians5 <- pedestrians %>%
  filter(TrajID == 5) %>%
  select(E,N) %>%
  as.matrix()

pedestrians6 <- pedestrians %>%
  filter(TrajID == 6) %>%
  select(E,N) %>%
  as.matrix()

```

then I compare trajectory 1 to trajectories 2-6 using different similarity 
measures from the package
```{r}
DTW = c(DTW(pedestrians1, pedestrians2), 
        DTW(pedestrians1, pedestrians3),
        DTW(pedestrians1, pedestrians4),
        DTW(pedestrians1, pedestrians5),
        DTW(pedestrians1, pedestrians6))

EditDist = c(EditDist(pedestrians1, pedestrians2),
             EditDist(pedestrians1, pedestrians3),
             EditDist(pedestrians1, pedestrians4),
             EditDist(pedestrians1, pedestrians5),
             EditDist(pedestrians1, pedestrians6))
                         
Frechet = c(Frechet(pedestrians1, pedestrians2),
            Frechet(pedestrians1, pedestrians3),
            Frechet(pedestrians1, pedestrians4),
            Frechet(pedestrians1, pedestrians5),
            Frechet(pedestrians1, pedestrians6))
                         
LCSS = c(LCSS(pedestrians1, pedestrians2, pointDistance = 20, pointSpacing = 0, errorMarg = 1),
         LCSS(pedestrians1, pedestrians3, pointDistance = 20, pointSpacing = 0, errorMarg = 1),
         LCSS(pedestrians1, pedestrians4, pointDistance = 20, pointSpacing = 0, errorMarg = 1),
         LCSS(pedestrians1, pedestrians5, pointDistance = 20, pointSpacing = 0, errorMarg = 1),
         LCSS(pedestrians1, pedestrians6, pointDistance = 20, pointSpacing = 0, errorMarg = 1))
```

The results are then putted together in a new data frame
```{r}
similarity_DF <- data.frame(ID = seq(1:5), DTW, EditDist, Frechet, LCSS)
```

Plotting the Results 
```{r}
# Long-Format
similarity_DF <- similarity_DF %>% 
  pivot_longer(-ID, names_to = "Method", values_to = "Value")

# Plot results
ggplot(similarity_DF, aes(as.factor(ID), Value, fill = as.factor(ID))) +
  geom_bar(stat = "identity") +
  facet_wrap(~Method, scales = "free_y") +
  theme(legend.position = "none") +
  labs(title = "Computed similarities using different measures \n between trajectory 1 and all the others",
       y = "Value\n", x = "\nComparison trajectory")
```




